Parsing Postfix log files

### 2007-10-18 doc/logparser.tex
I'm now using the figure environment for graphs, so they flow better in the paper.  The paper needs proof reading now, as there may be old and broken references.

@@

An important property of a parser is how execution time scales with input
size: does it scale linearly, polynomially, or exponentially?
Graph~\ref{execution time vs file size vs number lines} shows the execution
time in seconds, file size in MB and tens of thousands of lines per log
file per log file.  All three lines run roughly in parallel, giving a
visual impression that the algorithm scales linearly with input size.  This
impression is borne out by graph~\ref{execution time vs file size vs number
lines factor}, which plots the ratio of file size vs execution time and
number of lines vs execution time.  As the reader can see, the ratios are
quite tightly banded, showing that the algorithm scales linearly: the much
larger log files between points 60 and 70 on the X axis in
graph~\ref{execution time vs file size vs number lines} don't cause any
abnormality in the corresponding points in graph~\ref{execution time vs
file size vs number lines factor}.

@@

Graphs~\ref{normal vs shuffled vs reversed ordering factor} show the
percentage increase of execution times (mean and standard deviation are
shown after the graph).  Overall this provides a modest but worthwhile
performance increase.

@@

Perl compiles the original regex into an internal representation,
optimising the regex to improve the speed of matching, but this compilation
and optimisation takes CPU time.  Perl automatically caches static regexs,
but dynamic regexs need to be explicitly compiled and cached.
Graph~\ref{normal regex vs discard regex} shows execution times with and
without caching the regex.  Caching the compiled regexs is obviously far
more efficient; graph~\ref{normal regex vs disarded regex factor} shows the
percentage execution time increase when not caching each regex.

@@

@@

\subsubsection{Execution time vs file size vs number of lines.}
\label{execution time vs file size vs number lines}

\noindent
\includegraphics{build/plot-normal-filesize-numlines.ps}

@@

\subsubsection{Ratio of file size and number of lines to execution time}
\label{execution time vs file size vs number lines factor}

\noindent
\includegraphics{build/plot-normal-filesize-numlines-factor.ps}

@@

@@

\subsubsection{Percentage increase of shuffled + reversed over normal}
\label{normal vs shuffled vs reversed ordering factor}
\includegraphics{build/plot-normal-shuffle-factor.ps}
\newline
\includegraphics{build/plot-normal-reverse-factor.ps}

@@

The dips at logfiles 22 and 62--68 correspond to peaks in log file size in
graph~\ref{execution time vs file size vs number lines}, and peaks in
graph~\ref{execution time vs file size vs number lines factor} (where a
peak means that more lines are processed per second, i.e.\ performance is
better).  The explanation for this took some time to arrive at, but it
turns out to be reasonably simple.  The large log files in question were
caused by a mail forwarding loop, where the distribution of log lines is
quite different to normal, resulting in different performance
characteristics.

@@

The vast majority of log lines when a mail loop occurs are from Postfix
components which have a small number of rules associated with them, whereas
in general the most common log lines are logged by \textit{postfix/smtpd},
which also has the highest number of rules.  \textit{postfix/smtpd\/} log
lines are distributed across rules much more evenly that
\textit{postfix/qmgr}, \textit{postfix/local}, \textit{postfix/cleanup\/}
or \textit{postfix/pickup}, so the average number of rules required to
parse a \textit{postfix/smtpd\/} log line is much higher that the average
number required to parse other log lines.  Thus the average number of rules
required to parse a random line from logs containing a mail loop is much
lower than the average number required to parse a random log line from logs
without a mail loop, as can be seen in the peaks in graphs~\ref{execution
time vs file size vs number lines factor} and~\ref{normal vs shuffled vs
reversed ordering factor}.  The difference is more extreme in the reversed
ordering case because the largest number of rules to be consulted for the
commonly occurring log lines during a mail loop is far less than the
largest number of rules that are consulted for commonly occurring lines
outside of a mail loop.  The number of rules which needs to be consulted
when the ordering is shuffled varies between the optimum and nadir, and
the performance varies proportionally.

@@

\subsubsection{Regex: cached vs discarded}
\label{normal regex vs discard regex}
\includegraphics{build/plot-cached-discarded.ps}

\subsubsection{Regex: percentage execution time increase}
\label{normal regex vs disarded regex factor}
\includegraphics{build/plot-cached-discarded-factor.ps}

Two large dips can be seen in graph~\ref{normal regex vs disarded regex
factor} at log files 22 and 62--68, corresponding to the spikes in log file
size in graph~\ref{execution time vs file size vs number lines}.  The
reason for the anomalous log files has already been explained in
section~\ref{Why are there dips in the graphs?}.

@@

\noindent\includegraphics{build/plot-hits.ps}

As the graph above is quite difficult to read it has been separated into
three sections below: low, middle and high.

\noindent\includegraphics{build/plot-hits-low.ps}

It is apparent from the low graph that some rules have few or no hits;
those with zero hits are rules which were written to parse log files used
during development of the parser but not utilised in the test runs
performed for this document.

\noindent\includegraphics{build/plot-hits-middle.ps}

\noindent\includegraphics{build/plot-hits-high.ps}

>>>>

% \showgraph{filename}{caption}{label}
\newcommand{\showgraph}[3]{
    \begin{figure}[hbt!]
        \caption{#2}\label{#3}
        \includegraphics{#1}
    \end{figure}
}

@@

An important property of a parser is how execution time scales with input
size: does it scale linearly, polynomially, or exponentially?
Graph~\ref{execution time vs file size vs number of lines graph} shows the
execution time in seconds, file size in MB and tens of thousands of lines
per log file per log file.  All three lines run roughly in parallel, giving
a visual impression that the algorithm scales linearly with input size.
This impression is borne out by graph~\ref{execution time vs file size vs
number lines factor}, which plots the ratio of file size vs execution time
and number of lines vs execution time.  As the reader can see, the ratios
are quite tightly banded, showing that the algorithm scales linearly: the
much larger log files between points 60 and 70 on the X axis in
graph~\ref{execution time vs file size vs number of lines graph} don't
cause any abnormality in the corresponding points in graph~\ref{execution
time vs file size vs number lines factor}.

@@

Graphs~\ref{percentage increase of shuffled over normal}
and~\ref{percentage increase of reversed over normal} show the percentage
increase of execution times (mean and standard deviation are shown after
the graph).  Overall this provides a modest but worthwhile performance
increase.

@@

Perl compiles the original regex into an internal representation,
optimising the regex to improve the speed of matching, but this compilation
and optimisation takes CPU time.  Perl automatically caches static regexs,
but dynamic regexs need to be explicitly compiled and cached.
Graph~\ref{normal regex vs discard regex} shows execution times with and
without caching the regex.  Caching the compiled regexs is obviously far
more efficient; graph~\ref{normal regex vs discarded regex factor} shows
the percentage execution time increase when not caching each regex.

@@

\renewcommand{\figurename}{Graph}

@@

\showgraph{build/plot-normal-filesize-numlines.ps}{Execution time vs file
size vs number of lines}{execution time vs file size vs number of lines graph}

@@

\showgraph{build/plot-normal-filesize-numlines-factor.ps}{Ratio of file
size and number of lines to execution time}{execution time vs file size vs
number lines factor}

@@

\clearpage

@@

\showgraph{build/plot-normal-shuffle-factor.ps}{Percentage increase of
shuffled over normal}{percentage increase of shuffled over normal}

\showgraph{build/plot-normal-reverse-factor.ps}{Percentage increase of
reversed over normal}{percentage increase of reversed over normal}

@@

The dips at logfiles 22 and 62--68 correspond to peaks in log file size in
graph~\ref{execution time vs file size vs number of lines graph}, and peaks in
graph~\ref{execution time vs file size vs number lines factor} (where a
peak means that more lines are processed per second, i.e.\ performance is
better).  The explanation for this took some time to arrive at, but it
turns out to be reasonably simple.  The large log files in question were
caused by a mail forwarding loop, where the distribution of log lines is
quite different to normal, resulting in different performance
characteristics.

@@

The vast majority of log lines when a mail loop occurs are from Postfix
components which have a small number of rules associated with them, whereas
in general the most common log lines are logged by \textit{postfix/smtpd},
which also has the highest number of rules.  \textit{postfix/smtpd\/} log
lines are distributed across rules much more evenly that
\textit{postfix/qmgr}, \textit{postfix/local}, \textit{postfix/cleanup\/}
or \textit{postfix/pickup}, so the average number of rules required to
parse a \textit{postfix/smtpd\/} log line is much higher that the average
number required to parse other log lines.  Thus the average number of rules
required to parse a random line from logs containing a mail loop is much
lower than the average number required to parse a random log line from logs
without a mail loop, as can be seen in the peaks in graphs~\ref{execution
time vs file size vs number lines factor},~\ref{percentage increase of
shuffled over normal} and~\ref{percentage increase of reversed over
normal}.  The difference is more extreme in the reversed ordering case
because the largest number of rules to be consulted for the commonly
occurring log lines during a mail loop is far less than the largest number
of rules that are consulted for commonly occurring lines outside of a mail
loop.  The number of rules which needs to be consulted when the ordering is
shuffled varies between the optimum and nadir, and the performance varies
proportionally.

@@

\showgraph{build/plot-cached-discarded.ps}{Regex: cached vs
discarded}{normal regex vs discard regex}

\showgraph{build/plot-cached-discarded-factor.ps}{Regex: percentage
execution time increase}{normal regex vs discarded regex factor}

Two large dips can be seen in graph~\ref{normal regex vs discarded regex
factor} at log files 22 and 62--68, corresponding to the spikes in log file
size in graphs~\ref{execution time vs file size vs number of lines graph}
and~\ref{normal regex vs discard regex}.  The reason for the anomalous log
files has already been explained in section~\ref{Why are there dips in the
graphs?}.

@@

\showgraph{build/plot-hits.ps}{Hits per rule}{rule hits graph}

As graph~\ref{rule hits graph} is quite difficult to read it has been
separated into three sections: low, middle and high.

\showgraph{build/plot-hits-low.ps}{Hits per rule (low)}{hits per rule low}

It is apparent from the low graph (graph~\ref{hits per rule low}) that some
rules have few or no hits; those with zero hits are rules which were
written to parse log files used during development of the parser but not
utilised in the test runs performed for this document.

\showgraph{build/plot-hits-middle.ps}{Hits per rule (middle)}{hits per rule
middle}

\showgraph{build/plot-hits-high.ps}{Hits per rule (high)}{hits per rule
high}

\clearpage
